SODA POP : Robust online Partial Ordering Planning for real life applications

### 2015-11-22 article/article.tex
fixing comments and reducting the size

One of the applications that needs these improvements is plan
recognition with the particular use of off-the-shelf planners to infer
the pursued goal of an agent where online planning and resilience is
particularly important {[}5{]}. This method adds dummy actions that need
to be satisfied by modifying the goal to ensure that the observed plan
is selected by the planner. A system that is able to repair plan in real
time will ease such an application. Another application is
decision-making in dynamical environments. Indeed, having a plan that
details all steps and explicits the ones that are not possible, can help
decide of the line of action to take.

@@

A partial plan is valid if and only if it is consistent and if all the
flat plan that can be generated are valid.

@@

Partial Order Planning (POP) is a popular implementation of the general
PSP algorithm. It is proven to be sound and complete {[}6{]}. The
completeness of the algorithm guarantees that if the problem has a
solution it will be found by the algorithm. The soundness assures that
any answer from the algorithm is valid. POP refines a partial plan by
trying to fix its flaws.

@@

\begin{algorithm}\caption{Classical Partial Order Planning algorithm}\label{pop}\begin{algorithmic}[1]

@@

From that point the base algorithm is very similar for any
implementation of POP : using an agenda of flaws that is efficiently
updated after each refinement of the plan. A flaw is selected for
resolution and we use a non deterministic choice operator to pick a
resolver for the flaw. The resolver is inserted in the plan and we
recursively call the algorithm on the new plan. On failure we return to
the last non deterministic choice to pick another resolver. The
algorithm ends when the agenda is empty or when there is no more
resolver to pick for a given flaw.

@@

Some other closer works like {[}4{]} treats the problem of online
planning by removing big chunks of the partial plan by identifying
incorrect trees in the plan. This along with heuristic choice of
unrefinemment strategies. This causes a heavy replanning of the problem
even if only one action needed removal. This also lead to an exponential
number of plan to consider. This is a big problem when trying to adapt a
plan with minimal changes due to replanning.

@@

\begin{figure}[htbp]
\centering
\includegraphics{graphics/problem.pdf}
\caption{A simple problem that will be used throughout this
paper\label{fig:problem}}
\end{figure}

Before continuing, we present a simple example of classical POP
execution with the problem represented in figure \ref{fig:problem}. We
did not represent empty preconditions or effects to improve readability.
Here we have an initial state
\(I = \langle \emptyset , \{ 1, 2 \} \rangle\) and a goal
\(G = \langle \{ 3, 4, -5, 6 \}, \emptyset \rangle\) encoded as dummy
steps. We also introduce actions that are not steps yet but that are
provided by \(A\). The actions \(a\), \(b\) and \(c\) are normal actions
that are useful to achieve the goal. The action \(t\) is meant to be
threatening to the plan's integrity and will generate threats. We
introduce \(u\) and \(v\) as a toxic action, \(w\) as a dead-end action
and \(x\) as a contradictory action. These notions will be defined
\protect\hyperlink{defects}{later}.

@@

Of course the flaw selection mechanism of certain variant can prevent
that to happen in that case. But often flaw selection mechanisms are
more speed oriented and will do little if a toxic action seems to fit
better than a more coherent but complex one.

@@

In order to improve POP algorithms' resilience, online performance and
plan quality, we propose a set of auxiliary algorithms that provides POP
with a clean and efficiently populated initial plan. The complete
algorithm will be presented in \protect\hyperlink{soda}{the next
section} as a combination of all auxiliary algorithms and regular POP.

@@

\begin{algorithm}\caption{Proper plan generation algorithm for a given goal $g$}\label{properplan}\begin{algorithmic}[1]

@@

As in online planning goals can be known in advance, we add a new
mechanism that generates proper plans for goals. We take advantage of
the fact that this step can be done offline to improve performance for
online planning. This offline execution prevents us to access the
details of the initial state of the world as it will be defined at
runtime. We define for that the concept of \emph{participating action}.
An action \(a \in A\) participates in a goal \(G\) if and only if \(a\)
has an effect \(f\) that is needed to accomplish \(G\) or that is needed
to accomplish another participating action's preconditions. A proper
plan is a partial plan that contains all participating actions as steps
and causal links that bind them with the step they are participating in.
This proper plan is independent from the initial step because we might
not have the initial step at the time of the proper plan generation.

@@

On our example it illustrates really well the need for resilience. The
algorithm simply recursively chose the actions and end up with the
partial plan represented in figure \ref{fig:proper}. The partial plan
doesn't have an initial state (because of its offline nature). This plan
also shows several cycles and obvious problems. Moreover, the plan has
all the steps of the correct final plan and therefore remove the subgoal
search in POP if it uses this plan has input (after it has been cleaned
obviously).

This auxiliary algorithm is therefore used as a caching mechanism for
online planning. The algorithm starts to populate the proper plan with a
quick and incomplete backward chaining.

@@

\begin{algorithm}\caption{Defect resolution algorithm}\label{defectresolution}\begin{algorithmic}[1]

@@

When the POP algorithm is used to refine a given plan (that was not
generated with POP or that was altered), a set of new defects can be
present in it interfering in the resolution and sometimes making it
impossible to solve. We emphasize that these defects are not regular POP
flaws but new problems that classical POP can't solve. The aim of this
auxiliary algorithm is to clean the plans from such defects in order to
improve computational time, resilience and plan quality. It should be
noted that in some cases cleaning plans will increase the number of
flaws in the plan but will always improve the overall quality of it.

@@

From there it is obvious that plan quality will improve over POP since
the present algorithm is guaranteed to remove all defects in a plan.

@@

These defects are usually hypothesized out by regular models. They are
illegal use of partial plan representation and shouldn't happen under
regular circumstances. They may appear if the input is generated by an
approximate cognitive model that doesn't ensure consistency of the
output or by unchecked corrupted data. These defects will most of the
time simply break regular POP algorithms (from existing cycles) or at
least make the performances decrease significantly (from toxic actions).

@@

We use strongly connected component detection algorithm to detect
cycles. Upon cycle detection the algorithm can remove arbitrarily a link
in the cycle to break it. The most effective solution is to remove the
link that is the farthest from the goal traveling backward as it would
be that link that would have been last added in the regular POP
algorithm. But finding this is extremely time consuming and most of the
time the cycles are of one or two actions so we choose to keep that
algorithm simpler and faster.

@@

In a plan some actions can be illegal for POP. Those are the actions
that are contradictory. An action \(a\) is contradictory if and only if
\[\exists f \{f, \lnot f \} \in eff(a) \lor \{f, \lnot f \} \in pre(a)\]

@@

We remove only one of those effect or preconditions based on the usage
of the action in the rest of the plan. If none of those are used, we
choose to remove both. In our example in figure \ref{fig:problem} the
action \(x\) is one of these inconsistent actions with fluent \(2\) and
\(-2\) in its effects.

@@

These actions are those that have effects that are already in their
preconditions or empty effects. This can damage a plan as well as make
the execution of POP algorithm much longer than necessary. They are
defined as :
\[a | pre(a) \cap eff(a) \neq \emptyset \lor eff(a) = \emptyset\]

@@

This is fixed by removing the toxic fluents
(\(pre(a) \nsubseteq eff(a)\)) following \(eff(a) = eff(a)-pre(a)\). If
the effects gets empty the action is removed altogether from plan and
\(A\). For example in figure \ref{fig:problem} the actions \(u\) and
\(v\) are a toxic action as the fluent \(4\) is in the effects and
preconditions of \(v\). In this case these actions will be removed by
the algorithm as it doesn't have any other effects.

@@

The defects can be related to incorrect links. The first of which are
liar links : a link that doesn't reflect the preconditions or effect of
its source and target. We can note
\[a_i \xrightarrow{f} a_j | f \notin eff(a_i) \cap pre(a_j)\]

@@

This kind of defects is not as toxic as the illegal ones: they won't
make the plan unsolvable but they can still cause performance drops in
POP execution. These can appear more often in regular POP results as
they are not targeted by standard implementations.

@@

This defect can happen in POP generated plans to some extends. A
redundant link have a transitive equivalent of longer length. It means
that a link \(a_i \to a_j\) is redundant if and only if it exists
another path from \(a_i\) to \(a_j\) of length greater to \(1\).

@@

Since POP relies on those additional links, this part focus on removing
the ones that were generated for threat removal purpose to simplify the
plan. For example in the figure \ref{fig:pop} we can see that the
ordering constraint from \(a\) to \(t\) is redundant with the path
\(a \xrightarrow{5} c \to t\) in that it isn't needed to resolve the
threat. Therefore, the algorithm would remove it. This reduce the number
of edges in the plan and therefore simplify it.

@@

Causal links can be found to compete with one another. A competing link
\(a_i \xrightarrow{f} a_k\) competes with another link
\(a_j \xrightarrow{f} a_k\) if it provides the same fluent to the same
action.

@@

This cannot happen in classical POP algorithm so it is not handled by
it. The multiple links providing the same fluent to the same action is
making the plan more complex like redundant links but also can hide
useless actions. This is why they must be removed in a way that reveals
them and reduce the need for extra unnecessary actions in the plan that
would cause more flaws for POP to handle.

@@

In order to prune the least useful actions, we need to remove the least
interesting link. In order to elect the best one, we compare their
respective providing action.

@@

The usefulness of an action is the number of participating links
(outgoing causal links) divided by the number of needing links (incoming
causal links). If there are no needing links, or the action is either an
initial step or a goal the usefulness is \(+\infty\). The usefulness of
savior action is \(0\).

@@

We choose the link with the source action having the highest usefulness.
This makes replacing unnecessary savior actions and allows to therefore
reduce the usefulness of the least action in the hope it will become
orphan and pruned from the plan.

@@

@@

Actions can sometimes have no use in a plan as they don't contribute to
it. It is the case of orphans actions (actions without any links) and
action with no outgoing path to the goal (meaning that it doesn't
participate in the plan). That also concerns action that would become
orphan if another orphan action is removed.

@@

In order to fix this we derive the idea of the backward removal tree of
{[}4{]}. The idea is to remove an action if all its outgoing causal
links are leading to useless actions. We iterate over all actions until
no new useless action is discovered. All useless actions are then
removed from the plan.

\begin{algorithm}\caption{Orphan actions finding algorithm}\label{orphanaction_find}\begin{algorithmic}[1]

@@

@@

\begin{algorithm}\caption{Soft resolution healing algorithm}\label{softresolution}\begin{algorithmic}[1]

@@

@@

The concept of healer is made to target rogue flaws that caused total
failure. A healer is a resolver that is built around the savior of the
flaw in order to provide the missing fluents to it. The general formula
of a healer is the following : \[a_s \xrightarrow{p} a_n\] with \(a_s\)
being the savior of the flaw.

@@

Once the POP algorithm fails completely, the soft failing algorithm can
be invoked to heal the plan. It chooses the best plan
\(b | \forall p \in F, b \prec p\) to heal it. If two plans have the
same violation degree, the algorithm chooses one arbitrarily .

@@

After defining the way our algorithms work, we will focus on the
properties that can be achieved by combining them together.

\begin{algorithm}\caption{Combining all algorithms into SODA POP}\label{soda}\begin{algorithmic}[1]

@@

\begin{proof}[Convergence of POP over solvable problemes]

@@

Now for the more complex case of unsolvable problems we need to refer to
the way POP works.

\begin{proof}[Convergence of POP over unsolvable problemes]

POP algorithm will seek to solve flaws. Be \(flaws(p)\) the set of flaws
of a given partial plan.

The number of flaws is the number of subgoals plus the number of threats
\[flaws(p) = subgoals(p) \cup threats(p)\]

We consider the number of actions \(\#A\) as being finite. Therefore,
the number of steps in the plan is at worse \(\#A_p = \#A\). We also
assume that actions has a finite number of preconditions and effects
(since we don't use functions over fluents). This leads to :
\[\#subgoals(p) < \sum_{a \in A_p} \#pre(a) < \infty\]
\[\#threats(p) < \#L \le \sum_{a \in A_p} \#pre(a) < \infty\]

This means the number of all possible flaws is finite. As POP resolves
these flaws it will decrease their numbers and iterate over resolvers.
The number of resolver is \(\#subgoals(p) * \#A + \#threats(p) * 2\) and
is also finite. This means that the iteration will in the worst case be
of the number of resolvers before failing and this proves that the
algorithm terminates.

@@

The proper plan and defect fixing algorithms are obviously convergent.
The proper plan algorithm can't iterate more than the number of actions
(since duplicates are forbidden) and the defect resolution will fix the
finite numbers of defects present in the finite partial plan issued by
the proper plan algorithm.

POP is sound and converges. Therefore, if SODA converges it will return
a valid solution for \(P'\). In the same way we proved for POP, the new
healing process converges because it reduces the number of flaws in the
partial plan and since this number is finite, the algorithm converges.

@@

SODA POP algorithms can be used in dynamic online environments to allow
a robust way to replan an existing obsolete plan. The first thing to
execute prior to the runtime is the generation of the proper plans of
all the goals that will be considered.

For the first execution of the online planning the initial state is
added to the problem and the defect detection is applied with the rest
of the SODA POP algorithms. This will give the first plan.

On change of environment the previous plan is modified accordingly and
then fed to the defect detection and SODA algorithm again. This must be
done each time the plan is modified in order to actualize the plan. The
new plan will have all defects fixed. For example a savior action can be
removed if the initial step or a new action becomes competing with it
and actions with liar link to a removed fluent in the goal will become
probably an orphan.

>>>>

One of the applications that needs these improvements is plan
recognition with the particular use of off-the-shelf planners to infer
the pursued goal of an agent where online planning and resilience is
particularly important {[}5{]}. This method adds dummy actions that need
to be satisfied by modifying the goal to ensure that the observed plan
is selected by the planner. A system that is able to repair plan in real
time will ease such an application. Another application is
decision-making in dynamical environments. Indeed, having a plan that
details all steps and explicits the ones that are not possible, can help
decide of the plan of action to take.

@@

A partial plan is valid if and only if it is consistent and if all flat
plans that can be generated are valid.

@@

Partial Order Planning (POP) is a popular implementation of the general
PSP algorithm. It refines a partial plan by trying to fix its flaws and
is proven to be sound and complete {[}6{]}.

@@

\begin{algorithm}\caption{Classical Partial Order Planning}\label{pop}\begin{algorithmic}[1]

@@

Algorithm 1 presents the base algorithm for a planer in the plan space.
POP implementation uses an agenda of flaws that is efficiently updated
after each refinement of the plan. A flaw is selected for resolution and
we use a non deterministic choice operator to pick a resolver for the
flaw. The resolver is inserted in the plan and we recursively call the
algorithm on the new plan. On failure we return to the last non
deterministic choice to pick another resolver. The algorithm ends when
the agenda is empty or when there is no more resolver to pick for a
given flaw.

@@

Some other closer works like {[}4{]} treats the problem of online
planning by removing big chunks of the partial plan by identifying
incorrect trees in the plan. This along with a heuristic of choice of
unrefinemment strategies causes a heavy replanning of the problem even
if only one action needs removal. This also lead to an exponential
number of plan to consider. This is a big problem when trying to adapt a
plan with minimal changes due to replanning.

@@

Before continuing, we present a simple example of classical POP
execution with the following problem. We have an initial state
\(I = \langle \emptyset , \{ 1, 2 \} \rangle\) and a goal
\(G = \langle \{ 3, 4, -5, 6 \}, \emptyset \rangle\) encoded as dummy
steps. We also introduce actions that are not steps yet but that are
provided by \(A\). The actions \(a \langle \{1\}, \{3,5\} \rangle\),
\(b \langle \{2\}, \{4\} \rangle\) and
\(c \langle \{5\}, \{6\} \rangle\) are normal actions that are useful to
achieve the goal. The actions \(n \langle \{-8, 7\}, \{-7, 8\} \rangle\)
and \(l \langle \{-7, 8\}, \{-8, 7\} \rangle\) are looping actions that
are meant to cause cycles. The action
\(t \langle \{4\}, \{-5\} \rangle\) is meant to be threatening to the
plan's integrity and will generate threats. We introduce
\(u \langle \{4\}, \emptyset \rangle\) and
\(v \langle \{4\}, \{4\} \rangle\) as toxic actions,
\(w \langle \{9\}, \{3\} \rangle\) as a dead-end action and
\(x \langle \{7\}, \{-2,2\} \rangle\) as a contradictory action. These
notions will be defined \protect\hyperlink{defects}{later}.

@@

Of course the flaw selection mechanism of certain variant can prevent
that to happen in that case. But often flaw selection mechanisms are
more speed oriented and will do little if a toxic action seems to fit
better than a more coherent but complex one .

@@

In order to improve POP algorithms' resilience, online performance and
plan quality, we propose a set of auxiliary algorithms that provides POP
with a clean and efficiently populated initial plan. The complete
algorithm will be presented in \protect\hyperlink{soda}{the next
section} as a combination of all auxiliary algorithms and classical POP.

@@

\begin{algorithm}\caption{Proper plan generation for a given goal $g$}\label{properplan}\begin{algorithmic}[1]

@@

As in online planning goals can be known in advance, we propose a new
mechanism that generates proper plans for goals. We take advantage of
the fact that this step can be done offline to improve the performance
of online planning. This offline execution prevents us to access the
details of the initial state of the world as it will be defined at
runtime. We define for that the concept of \emph{participating action}.
An action \(a \in A\) participates in a goal \(G\) if and only if \(a\)
has an effect \(f\) that is needed to accomplish \(G\) or that is needed
to accomplish another participating action's preconditions. A proper
plan is a partial plan that contains all participating actions as steps
and causal links that bind them with the step they are participating in.
This proper plan is independent from the initial step because we might
not have the initial step at the time of the proper plan generation.

@@

We define the concept of \emph{participating action}. An action
\(a \in A\) participates in a goal \(G\) if and only if \(a\) has an
effect \(f\) that is needed to accomplish \(G\) or to accomplish another
participating action's preconditions. A proper plan is a partial plan
that contains all participating actions as steps and causal links that
bind them with the step they are participating in. Proper plan
generation, detailed in Algorithm 2, recursively chooses all
participating actions. It simply populate the proper plan with a quick
and incomplete backward chaining.

The partial plan doesn't have an initial state (because of its offline
nature). This auxiliary algorithm is therefore used as a caching
mechanism for online planning. The algorithm starts to populate the
proper plan with a quick and incomplete backward chaining.

Once applied on the example of figure \textbf{???}, proper plan
generation returns the partial plan presented in figure
\ref{fig:proper}. This partial plan doesn't have initial state because
of its offline nature. It also shows several cycles and obvious
problems. However, it has all the steps of the correct final plan.

This algorithm helps the POP algorithm by prefetching all the actions
subgoals might need but also need to be cleaned first to be helpful.

@@

\begin{algorithm}\caption{Defect resolution}\label{defectresolution}\begin{algorithmic}[1]

@@

When the POP algorithm is used to refine a given plan (that was not
generated with POP or that was altered), a set of new defects can be
present in it interfering in the resolution and sometimes making it
impossible to solve. We emphasize that these \emph{defects are not
regular POP flaws} but new problems that classical POP can't solve. The
aim of this auxiliary algorithm is to clean the plans from such defects
in order to improve computational time, resilience and plan quality. It
should be noted that in some cases cleaning plans will increase the
number of flaws in the plan but will always improve the overall quality
of it.

@@

From there it is obvious that plan quality will improve over POP since
our present algorithm is guaranteed to remove all defects in a plan.

@@

These defects are usually hypothesized out by regular models. They are
illegal use of partial plan representation and shouldn't happen under
regular circumstances. They may appear if the input is generated by an
approximate cognitive model that doesn't ensure consistency of the
output or by unchecked corrupted data. These defects will most of the
time simply break regular POP algorithms or at least make the
performances decrease significantly Illegal defects and solutions to fix
them are presented in the following section.

@@

We use strongly connected component detection algorithm to detect
cycles. Upon cycle detection the algorithm can remove arbitrarily a link
in the cycle to break it.

In a plan some actions can be illegal for POP. Those are the actions
that are inconsistent.

@@

An action \(a\) is contradictory if and only if
\[\exists f \{f, \lnot f \} \in eff(a) \lor \{f, \lnot f \} \in pre(a)\]

@@

We remove only one of those effect or preconditions based on the usage
of the action in the rest of the plan. If none of those are used, we
choose to remove both. In our example in figure \textbf{???} the action
\(x\) is one of these inconsistent actions with fluent \(2\) and \(-2\)
in its effects.

@@

Toxic actions are those that have effects that are already in their
preconditions or empty effects. This can damage a plan as well as make
the execution of POP algorithm much longer than necessary. They are
defined as :
\[a | pre(a) \cap eff(a) \neq \emptyset \lor eff(a) = \emptyset\]

@@

They can damage a plan as well as make the execution of POP algorithm
much longer than necessary. This is fixed by removing the toxic fluents
(\(pre(a) \nsubseteq eff(a)\)) following \(eff(a) = eff(a)-pre(a)\). If
the effects becomes empty, the action is removed altogether from plan
and \(A\). For example in figure \textbf{???} the actions \(u\) and
\(v\) are toxic as the fluent \(4\) is in the effects and preconditions
of \(v\), and \(u\) has empty effects. In this case these actions will
be removed by the algorithm as they don't have any other effects.

The defects can be related to incorrect links. The first of which are
liar links.

@@

A liar link is a link that doesn't reflect the preconditions or effect
of its source and target. We can note:
\[a_i \xrightarrow{f} a_j | f \notin eff(a_i) \cap pre(a_j)\]

@@

This kind of defects is not as problematic as the illegal ones: they
won't make the plan unsolvable but they can still cause performance and
quallity drops. These defects can happen in POP generated plans to some
extends. Interference defects and solutions to fix them are presented in
the following.

@@

A redundant link have a transitive equivalent of longer length. It means
that a link \(a_i \to a_j\) is redundant if and only if it exists
another path from \(a_i\) to \(a_j\) of length greater to \(1\).

@@

Since POP relies on those additional links, this part focus on removing
the ones that were generated for threat removal purpose to simplify the
plan. For example in figure \ref{fig:pop} we can see that the ordering
constraint from \(a\) to \(t\) is redundant with the path
\(a \xrightarrow{5} c \to t\) in that it isn't needed to resolve the
threat. Therefore, the algorithm would remove it. This reduces the
number of edges in the plan and therefore simplifies it.

Causal links can be found to compete with one another.

@@

A competing link \(a_i \xrightarrow{f} a_k\) competes with another link
\(a_j \xrightarrow{f} a_k\) if it provides the same fluent to the same
action.

@@

The main difference with redundant links is that competing links are not
related to threat resolution generated ordering constraint but confronts
two causal links that can provide the same fluent. Another difference is
that it cannot happen in classical POP algorithm. This kind of defect is
making the plan more complex and can also hide useless actions which
would potentially cause more flaws to handle than necessary.

@@

In order to remove the least interesting link in competing causal links,
we need to compare their respective providing action to chose the most
useful ones.

@@

The usefulness of an action is the number of participating links
(outgoing causal links) divided by the number of needing links (incoming
causal links). If there are no needing links, the usefulness is simmply
the number of participating links.

In the special cases if the action is either an initial step or a goal,
the usefulness is \(+\infty\). The usefulness of savior action is \(0\).

@@

We choose the link with the source action having the highest usefulness.
This replaces unnecessary savior actions and allows reducing the
usefulness of the least action in order to eventually make it an orphan
and prune it from the plan.

@@

On the left example of figure \ref{fig:competing}, the action \(a_j\)
participates much more than the action \(a_i\) and therefore the link to
be removed would be \(a_i \xrightarrow{f} a_k\). On the right example
the actions don't have different outgoing links but the action \(a_j\)
is here much needier than its competitor. Therefore, the link to be
removed is the link \(a_j \xrightarrow{f} a_k\).

Actions can sometimes have no use in a plan as they don't contribute to
it.

@@

Orphans actions are actions without any links or action with no outgoing
path to the goal (meaning that it doesn't participate in the plan). That
also concerns action that would become orphan if another orphan action
is removed.

@@

\begin{algorithm}\caption{Orphan actions finding}\label{orphanaction_find}\begin{algorithmic}[1]

@@

In order to fix this we derive the idea of the backward removal tree of
{[}4{]}. The algorithm, detailed in Algorithm 4, removes an action if
all its outgoing causal links are leading to useless actions. It
iterates over all actions until no new useless action is discovered. All
useless actions are then removed from the plan.

@@

\begin{algorithm}\caption{Soft resolution healing}\label{softresolution}\begin{algorithmic}[1]

@@

The concept of healer is made to target rogue flaws that caused total
failure.

@@

A healer is a resolver that is built around the savior of the flaw in
order to provide the missing fluents to it. The general formula of a
healer is the following : \[a_s \xrightarrow{p} a_n\] with \(a_s\) being
the savior of the flaw.

@@

Once the POP algorithm fails completely, the soft failing algorithm can
be invoked to heal the plan. It chooses the best plan
\(b | \forall p \in F, b \prec p\) to heal it. If two plans have the
same violation degree, the algorithm chooses the first one to have
happened.

@@

The combination of classical POP and all the auxiliary algorithms
presented previously gives SODA POP algorithm detailed in Algorithm 6.
In this section we will focus on the properties of SODA POP.

\begin{algorithm}\caption{SODA POP}\label{soda}\begin{algorithmic}[1]

@@

\begin{proof}

@@

Be \(flaws(p)\) the set of flaws of a given partial plan. The number of
flaws is the number of subgoals plus the number of threats
\(flaws(p) = subgoals(p) \cup threats(p)\)

We consider the number of actions \(\#A\) as being finite. Therefore,
the number of steps in the plan is at worse \(\#A_p = \#A\). We also
assume that actions has a finite number of preconditions and effects
(since we don't use functions over fluents).

This leads to \(\#subgoals(p) < \sum_{a \in A_p} \#pre(a) < \infty\) and
\(\#threats(p) < \#L \le \sum_{a \in A_p} \#pre(a) < \infty\)

This means the number of all possible flaws is finite. As POP resolves
these flaws it will decrease their numbers and iterate over resolvers.
The number of resolver is \(\#subgoals(p) * \#A + \#threats(p) * 2\) and
is also finite. This means that the iteration will in the worst case be
of the number of resolvers before failing and this proves that the
algorithm terminates. \qedhere

@@

The proper plan and defect fixing algorithms are obviously convergent.
The proper plan algorithm cannot iterate more than the number of actions
(since duplicates are forbidden) and the defect resolution will fix the
finite numbers of defects present in the finite partial plan issued by
the proper plan algorithm.

POP is sound and converges. Therefore, if SODA POP converges it will
return a valid solution for \(P'\). In the same way we proved for POP,
the new healing process converges because it reduces the number of flaws
in the partial plan and since this number is finite, the algorithm
converges.\qedhere

@@

SODA POP algorithm can be used in dynamic online environments to allow a
robust way to replan an existing obsolete plan. The first thing to
execute prior to the runtime is the generation of the proper plans for
all the goals that will be considered.

For the first execution of the online planning the initial state is
added to the problem and the defect detection is applied with the rest
of the SODA POP algorithm. This will give the first plan.

For every change of the environment the previous plan is modified
accordingly and then fed to the defect detection and SODA algorithm
again. This must be done each time the plan is modified in order to
actualize the plan. The new plan will have all defects fixed. For
example a savior action can be removed if the initial step or a new
action becomes competing with it and actions with liar link to a removed
fluent in the goal will become probably an orphan.
